One-loop PDF of the average Ising magnetization in 4 dimensions.

### 2020-03-08 paper/paper.tex
Paper checkpoint

\documentclass[11pt,a4paper]{article}
\usepackage{jheppub}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\newcommand{\expval}[1]{\left\langle#1\right\rangle}
\newcommand{\dd}{\mathrm{d}}

\title{One-loop effective action for the average Ising magnetization on compact manifolds}
\author{Andrea Allais}
\emailAdd{allais.andrea@gmail.com}

@@

The average magnetization $\bar{\phi} = \frac{1}{V} \sum_{i} \sigma_i$ of the
Ising model is the simplest observable in Monte Carlo simulations. For large
system size, the probability distribution of $\bar{\phi}$ changes qualitatively
between the two phases of the model. In the symmetric phase, if the
correlation length $\xi$ is large in lattice units, but small compared to the
system size $R$, the distribution is well approximated by a unimodal Gaussian
centered at 0. In the broken symmetry phase, still for $1 \ll \xi \ll R$, the
distribution is well approximated by a bimodal superposition of two 
Gaussians centered at non-zero values $\pm \bar{\phi}_0$.

Near the critical point $1 \ll R \ll \xi$ the probability distribution is a
non-trivial function of $\bar{\phi}$, at least in two and three dimensions.  It
is not immediately clear if the same is true in four dimensions as well.  Since
the $\phi^4$ field theory that describes the Ising critical point in four
dimensions becomes weakly coupled in the infrared, one may expect the
probability distribution to be a Gaussian at the critical point as well.

To clarify this issue, we compute the one-loop effective action, or log
probability, for $\bar{\phi}$ in $4$ dimensions, and show that for
very large system size, the critical probability distribution is not Gaussian,
but rather of the form $p(\bar{\phi})\sim \exp(-a \bar{\phi}^4)$. We find that
corrections to this form vanish very slowly with increasing system size, like
$1/\log R$.

Thus we consider the $\phi^4$ field theory:

\begin{equation}
  S[\phi] = \int_{\mathcal{M}}\left(
    \frac{1}{2} \phi\left(\Delta + r_0\right)\phi +
    \frac{1}{2} \xi_0 \phi^2 \mathcal{R} +
    \frac{1}{24} u_0 \phi^4\right)
\end{equation}

on both the 4-torus $\mathcal{M} = S_1^4$ or the 4-sphere $\mathcal{M} = S_4$.
Here $\mathcal R$ denotes the scalar curvature. We define the effective action:

\begin{equation}
  e^{-S_{\mathrm{eff}}(\bar\phi)} = \int \left[\mathcal{D}\phi\right]
    \delta\left(\bar{\phi} - \frac{1}{V}\int_{\mathcal{M}}\phi\right)
    e^{-S[\phi]}\,,
\end{equation}

and we obtain:

\begin{equation}
\label{eq:effective_action}
\begin{split}
  \frac{S_{\mathrm{eff}}(\bar{\phi})}{V} = 
  &\frac{1}{2} \frac{\bar{\phi}^2}{R^2} \left(
    r R^2 + \eta + \frac{g}{6} f_1\left(r R^2 + \eta\right)
    + O\left(g^2\right)\right) + \\
  & \frac{2\pi^2}{9}g\bar{\phi}^4\left(
    1 - \frac{g}{2}f_2\left(r R^2 + \eta\right)
    + O\left(g^2\right)\right) + \\
  & \frac{8\pi^4}{81} g^3 \bar{\phi}^6 R^2\left(
    f_3\left(r R^2 + \eta\right)
    + O\left(g\right)\right) + O\left(g^4\bar{\phi}^8\right)\,,
\end{split}
\end{equation}

where $R$ is the radius of either manifold, and the coefficients
$(r,\,\eta,\,g)$ are the renormalized counterparts to $(r_0,\,\xi,\,u_0)$. An
explicit expression for the functions $f_1$, $f_2$ and $f_3$ is reported in the
appendix (\ref{eq:f_functions}), and they are plotted in
fig.~\ref{fig:f_functions}.

\begin{figure}
\begin{center}
\includegraphics{f_functions.png}
\end{center}
\caption{\label{fig:f_functions} The functions $f_i$}
\end{figure}

This is our central result, and we will now describe it at lenght, beginning
with its domain of validity and its dependence on the couplings on a manifold
of fixed radius $R$.

The coupling $\eta$ is the counterpart to the curvature coupling
$\xi_0$, and hence is exactly zero for the torus, whereas it is a free,
dimensionless parameter for the sphere.

The dimension-2 coupling $r$ controls the cross-over between the symmetric and
broken-symmetry phases. The expression (\ref{eq:effective_action}) is valid for
$r R^2 + \eta \gtrsim -1$. It becomes invalid for lower values of $r$ because
the condensation of higher modes of the laplacian invalidates the perturbative
expansion. This becomes manifest in a divergence of the functions $f_i$ as
their argument approaches -1 for the torus, or  -4 for the sphere.  On the
other hand, $f_1$, $f_2$, $f_3$ remain finite as their argument approaches
positive infinity, and hence for large $r$ the probability distribution of
$\bar{\phi}$ is a Gaussian:

\begin{equation}
  \frac{S_{\mathrm{eff}}(\bar{\phi})}{V} \sim \frac{1}{2} r \bar{\phi}^2
  \quad \text{for} \quad
  rR^2 \gg 1\,.
\end{equation}

@@

In general,
they depend on a renormalization scale $\mu$, and it is possible to derive an
expression for $S_{\mathrm{eff}}$ evaluated at a generic value of $\mu$. Here,
for simplicity we choose to evaluate (\ref{eq:effective_action}) at the scale
$\mu^2 = r + p^2$. This value of $\mu$ is optimal for the reliability of
perturbation theory, because it avoids the emergence of large logarithms.

  As is well
known, in $4 - \varepsilon$ dimensions it satisfies the renormalization group
equation:

\begin{equation}
  \mu \frac{\dd g}{\dd \mu} = -\varepsilon g + g^2\,.
\end{equation}

The Wilson-Fisher fixed point is attained at $g_{\star} = \varepsilon$, and
substituting $g \to \varepsilon$ in (\ref{eq:effective_action}) and dropping
higher order terms in $\varepsilon$ yields the effective action at the fixed
point to leading order in $\epsilon$.

The coupling $r$ satisfies the renormalization group equation:
\begin{equation}
  \mu \frac{\dd r}{\dd \mu} = \frac{g}{3}r
\end{equation}

$r$, $\eta$, and $g$ are renormalized couplings
that satisfy the renormalization group equations:

\begin{align}
  &\mu \frac{\dd g}{\dd \mu} = -\varepsilon g + g^2 \\
  &\mu \frac{\dd r}{\dd \mu} = \frac{g}{3}r \\
  &\mu \frac{\dd \eta}{\dd \mu} = \frac{g}{3}(\eta - 2)
\end{align}

@@

While a general expression for the effective action at a generic scale $\mu$ is
available, for simplicity (\ref{eq:effective_action}) is evaluated at $\mu^2
= r + p^2$. This value of $\mu$ is optimal for the reliability of perturbation
theory, because it minimizes the effect of the so-called large logs.

An explicit expression for the functions $f_1$, $f_2$, $f_3$ is reported in the
appendix. As shown in fig.~\ref{fig:f_functions}, $f_i(x)$ is of order 1 in for
$x \gtrsim -1/2$, $f_1$ approaches a constant as $x \to \infty$, whereas
$f_2$ and $f_3$ approach 0. They all diverge as $x$ approaches $-1$.
\section{Appendix}

\begin{equation}
\label{eq:f_functions}
f_1 = 
\end{equation}
\end{document}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage{jheppub}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{tikz}
\newcommand{\expval}[1]{\left\langle#1\right\rangle}
\newcommand{\dd}{\mathrm{d}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}

\title{One-loop PDF of the average Ising magnetization in 4 dimensions.}
\author{Andrea Allais}
\emailAdd{allais.andrea@gmail.com}

@@

The average magnetization $\bar{\phi} = \frac{1}{V} \sum_{i} \sigma_i$ of the
Ising model is the simplest observable in Monte Carlo simulations. For large
system size, the probability distribution of $\bar{\phi}$ changes qualitatively
between the two phases of the model. In the symmetric phase, if the correlation
length $\xi$ is large in lattice units, but small compared to the system size
$L$, the distribution is well approximated by a normal distribution centered at
0. In the broken symmetry phase, still for $1 \ll \xi \ll L$, the distribution
is well approximated by a bimodal superposition of two normals centered at
non-zero values $\pm \bar{\phi}_0$.

Near the critical point $1 \ll L \ll \xi$ the probability distribution of
$\bar{\phi}$ is a non-trivial function of $\bar{\phi}$, at least in two and
three dimensions.  It is not immediately clear if the same is true in four
dimensions as well.  Since the $\phi^4$ field theory that describes the Ising
critical point in four dimensions becomes weakly coupled in the infrared, one
may expect the probability distribution to be a single zero-centered normal at
the critical point as well.

To clarify this issue, we compute the log pdf of $\bar{\phi}$ at one-loop in
$4$ dimensions, and show that for very large system size, the critical
probability distribution is not normal, but rather of the form
$-\log p(\bar{\phi})\sim f(L) \bar{\phi}^4$. We find that corrections to this
form vanish very slowly with increasing system size, like $1/\log L$.

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{binder_cumulant.png}
\end{center}
\caption{\label{fig:binder_cumulant} Blah Blah}
\end{figure}

Thus we consider the $\phi^4$ field theory with action:
\begin{equation}
  S[\phi] \equiv \int \dd^4 x\ \left(
    \frac{1}{2} \phi\left(\Delta + r_0\right)\phi +
    \frac{1}{24} u_0 \phi^4\right)
\end{equation}
defined on a 4-torus of radius $R$, i.e. $x_i\sim x_i + 2\pi R$. In this field
theory context, the pdf of the average magnetization is defined as:
\begin{equation}
    \label{eq:pdf_definition}
    p(\bar\phi) \equiv \frac{1}{Z} \int \left[\mathcal{D}\phi\right]
    \delta\left(\bar{\phi} - \frac{1}{V}\int\dd^4 x\ \phi\right)
    e^{-S[\phi]}\,,
\end{equation}
and at one loop it evaluates to:
\begin{equation}
\label{eq:log_pdf}
\begin{split}
    -\log p(\bar{\phi}) = \mathrm{const.} +  V\Bigg(
  &\frac{1}{2} \frac{\bar{\phi}^2}{R^2} \left(
    r R^2 + \frac{g}{6} \left(f_1\left(r R^2\right) - 1\right)
    + O\left(g^2\right)\right) + \\
  & \frac{2\pi^2}{9}g\bar{\phi}^4\left(
    1 - \frac{g}{2}f_2\left(r R^2\right)
    + O\left(g^2\right)\right) + \\
  & \frac{8\pi^4}{81} g^3 \bar{\phi}^6 R^2\left(
    f_3\left(r R^2\right)
    + O\left(g\right)\right) + O\left(g^4\bar{\phi}^8\right) \Bigg)\,,
\end{split}
\end{equation}
where the coefficients $r$ and $g$ are the renormalized counterparts to $r_0$
and $u_0$. An explicit expression for the functions $f_1$, $f_2$ and $f_3$ is
reported in the appendix (\ref{eq:f_functions}), and they are plotted in
fig.~\ref{fig:f_functions}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{f_functions.png}
\end{center}
\caption{\label{fig:f_functions} The functions $f_i$}
\end{figure}

Eq. (\ref{eq:log_pdf}) is our central result, and we now describe it
at lenght. We first illustrate its domain of validity and its dependence on the
renormalized couplings $r$, $g$ on a torus of fixed radius $R$, then describe
its dependence on $R$ at fixed bare couplings $r_0$, $u_0$.

The dimension-2 coupling $r$ controls the cross-over between the symmetric and
broken-symmetry phases. The expression (\ref{eq:log_pdf}) is valid for
all positive values of $r$, and also for negative values, provided that $rR^2
\gtrsim -1/2$.  For lower values of $r$, additional modes of the laplacian
condense beside the zero-mode, and this invalidates the perturbative expansion.
. This becomes manifest in a divergence of the functions $f_i$ as their
argument approaches -1.  On the other hand, $f_1$, $f_2$, $f_3$ all go to zero
as their argument approaches positive infinity, and hence for large $r$ the
probability distribution of $\bar{\phi}$ is a zero-centered normal:
\begin{equation}
    p(\bar{\phi}) \sim N e^{- \frac{1}{2} V r \bar{\phi}^2}
  \quad \text{for} \quad
  rR^2 \gg 1\,.
\end{equation}

@@

In order to describe how $p(\bar{\phi})$ depends on $R$ at fixed bare couplings
$r_0$, $u_0$, it is necessary to account for renormalization effects. In
general, the renormalized couplings $r$, $g$ depend on a renormalization scale
$\mu$, and it is possible to derive an expression for $p(\bar{\phi})$ evaluated
at a generic value of $\mu$. However, for simplicity,
(\ref{eq:log_pdf}) is evaluated at the scale $\mu^2 = r + R^{-2}$.
This choice of $\mu$ is optimal for the reliability of perturbation theory,
because it avoids the emergence of large logarithms over the widest possible
range of parameters.

The renormalization group equations for $r$ and $g$ are:
\begin{equation}
  \label{eq:renormalization_group}
  \mu \frac{\dd g}{\dd \mu} = g^2\,,\quad
  \mu \frac{\dd r}{\dd \mu} = \frac{1}{3}g r\,.
\end{equation}

These can be integrated and combined with the condition $\mu^2 = r + R^{-2}$ to
obtain a system of equations\footnote{There are of course many alternative,
arguably simpler, solutions that differ by subleading orders in an expansion in
$g(R_1)$. The one displayed here is the exact solution to
(\ref{eq:renormalization_group}).} connecting the renormalized couplings at two
different values of $R$:
\begin{equation}
  \frac{g(R_1)}{g(R_2)} = \frac{r(R_1)^3}{r(R_2)^3} = 
  1 - \frac{1}{2} g(R_1)\log\frac{r(R_2) + R_2^{-2}}{r(R_1) + R_1^{-2}}\,.
\end{equation}

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{renormalization_group.png}
\end{center}
    \caption{\label{fig:renormalization_group} A few solutions to
    (\ref{eq:renormalization_group}) with $g(R_1) = 0.45$.  In the left plot,
    the shaded region shows where the perturbative expansion becomes invalid.
    Similarly in the right plot, the lines become dotted oustside of the
    perturbative region. Note how the sign of $r$ is preserved, $r = 0$ is a
    solution, and $g$ goes to $0$ for $R_2 \gg R_1$.}
\end{figure}

A few solutions to these equations are shown in
fig.~\ref{fig:renormalization_group}. Within the perturbative regime $g \ll 1$,
$r R^2 \gtrsim -1/2$, the coupling $r$ varies very little with $R$, and, as is
clear from the differential form (\ref{eq:renormalization_group}), the sign of
$r$ is always preserved. Thus we conclude that the critical point is at $r
= 0$, and the symmetric phase is realized for $r > 0$.

At the critical point, the renormalized coupling $g$ follows the simplified renormalization group equation:
\begin{equation}
    \frac{1}{g(R_2)} - \frac{1}{g(R_1)} = \log\frac{R_2}{R_1}\,,
\end{equation}
and hence, as system size grows, the renormalized coupling $g$ goes to zero.
In this regime, the quartic term in (\ref{eq:log_pdf}) dominates all the
others. This is perhaps most evident if the PDF is expressed in terms of the
rescaled quantity $\bar{\varphi} = g^{\frac{1}{4}} \bar{\phi}$, whose
variance remains finite as $g\to0$.

Thus we conclude that, for sufficiently large system size: 
\begin{equation}
    \label{eq:log_pdf_critical_point}
    -\log p(\phi) \sim \mathrm{const} + \frac{2\pi^2}{9} V g\bar{\phi}^4\,,
\end{equation}
with subleading terms vanishing like $1 / \log(R)$.

In Monte Carlo simulations, the qualitative behavior of the distribution of the
magnetization is often characterized by measuring the so-called Binder
cumulant:
\begin{equation}
    U = 1 - \frac{\left\langle \bar \phi^4 \right \rangle}
    {3\left\langle \bar \phi^2 \right \rangle^2},
\end{equation}
which is constructed to be independent of the overall scale of $\bar{\phi}$,
and to be zero if $\bar{\phi}$ is normally distributed. From
(\ref{eq:log_pdf_critical_point}) we conclude that, on a 4-Torus, at the
critical point\footnote{ Incidentally, we note that this result is not in
contraddiction with the claim in the original paper that the critical Binder
cumulant in 4 dimensions is zero, because in the original paper the cumulant is
not computed from the total magnetization, but rather from the average
magnetization of blocks of spins much smaller than the overall system size.}:
\begin{equation}
    \label{eq:critical_binder_cumulant}
    U = 1 - \frac{4\,\Gamma\left(\frac{5}{4}\right)^2}
    {3\,\Gamma\left(\frac{3}{4}\right)^2} 
    = 0.27052\ldots
\end{equation}

In fig.~\ref{fig:binder_cumulant}, we show a comparison of the Binder cumulant
computed from (\ref{eq:log_pdf}) near the critical point, and the results of
Monte Carlo simulation of the Ising model.

\section{Derivation}

We now describe briefly how the result (\ref{eq:log_pdf}) is obtained. First of
all we observe that the average magnetization $\bar{\phi}$ is proportional to
the zero momentum mode of the field $\phi$:
\begin{equation}
    \phi(x) = \frac{1}{V} \sum_{n \in \mathbb{Z}^4} \phi_{n} 
    e^{i \frac{n\cdot x}{R}}\,;\quad \bar{\phi} = \frac{\phi_{0}}{V}\,.
\end{equation}

Thus, because of the delta function in (\ref{eq:pdf_definition}), the component
$\phi_{0}$ becomes an external field, whereas the other components $\phi_{n}$
with are still part of the functional integral. Separating the zero momentum
mode from the other modes in the action yields:
\begin{equation}
\begin{split}
    S(\bar{\phi}, \phi) =& 
    V \left(\frac{1}{2} r_0 \bar{\phi}^2 + \frac{u_0}{24} \bar{\phi}^4\right) +
    \frac{u_0}{4V} \bar{\phi}^2 \sum_{n} \phi_{n} \phi_{-n} + \\
    &\frac{u_0}{6V^2} \bar{\phi} 
        \sum_{n_1, n_2} \phi_{n_1} \phi_{n_2} \phi_{-n_1 -n_2} + \\
    &\frac{u_0}{24V^3} \sum_{n_1, n_2, n_3} \phi_{n_1} \phi_{n_2} 
        \phi_{n_3} \phi_{-n_1 -n_2 -n_3}\,.
\end{split}
\end{equation}

We also introduce renormalized couplings
\begin{equation}
    u_0 = u(1 + u\delta u)\,;\quad r_0 = r(1 + u \delta r)\,,
\end{equation}
and we obtain the following edges and vertices in the diagrammatic expansion:
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[dashed, thick] (-1, -1) -- (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above right] at (1, 1) {$n_1$};
    \node[below right] at (1, -1) {$n_2$};
\end{tikzpicture}
    -\frac{u\bar{\phi}^2}{4V}\delta_{n_1 + n_2}
    \quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[dashed, thick] (-1, -1) -- (0, 0);
    \draw[thick] (0, 0) -- (-1, 1);
    \draw[thick] (+1, -1) -- (0, 0) -- (+1, 1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
\end{tikzpicture}
    -\frac{u\bar{\phi}}{6V^2}\delta_{n_1 + n_2 + n_3}
\end{displaymath}
\begin{displaymath}
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, -1) -- (1, 1);
    \draw[thick] (-1, 1) -- ( 1, -1);
    \node[above left] at (-1, 1) {$n_1$};
    \node[above right] at (1, 1) {$n_2$};
    \node[below right] at (1, -1) {$n_3$};
    \node[below left] at (-1, -1) {$n_4$};
\end{tikzpicture}
    -\frac{u}{24V^3}\delta_{n_1 + n_2 + n_3 + n_4}
\quad\quad
\begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
    \draw[thick] (-1, 0) -- (1, 0);
    \node[left] at (-1, 0) {$n_1$};
    \node[right] at (1, 0) {$n_2$};
\end{tikzpicture}
    \quad
    \frac{V}{n_1^2R^{-2} + r}\delta_{n_1 + n_2}
\end{displaymath}
plus additional vertices associated with the counterterms $\delta r$, $\delta
u$ which we do not list for brevity.

The logarithm of the probability distribution $p(\bar{\phi}$ is the sum of all connected diagrams:
\begin{displaymath}
\begin{split}
    \log p(\bar{\phi}) = &\,
    \mathrm{const} -V \left(\frac{1}{2} r \bar{\phi}^2 + 
    \frac{u}{24} \bar{\phi}^4\right)\\
&+\ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1, -0.5) -- (0, 0) -- (1, -0.5);
    \draw[thick] (0, 0) .. controls (-1.5, 1.3) and (1.5, 1.3) .. (0, 0);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1.75, 0) -- (-0.75, 0);
    \draw[thick] (-0.75, 0) -- (0.75, 0);
    \draw[thick, dashed] (1.75, 0) -- (0.75, 0);
    \draw[thick] (0, 0) circle (0.75cm);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1, -0.5) -- (0, 0) -- (1, -0.5);
    \draw[thick] (0, 0) .. controls (0.6, 0.5).. (0, 1) .. 
    controls (-1.5, 2.3) and (1.5, 2.3) .. (0, 1) .. 
    controls (-0.6, 0.5) .. (0, 0) ;
\end{tikzpicture}
 \ + \ \cdots \\
&+\ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1.75, -0.75) -- (-1, 0) -- (-1.75, 0.75);
    \draw[thick, dashed] (1.75, -0.75) -- (1, 0) -- (1.75, 0.75);
    \draw[thick] (-1, 0) .. controls (-0.2, 0.8) and (0.2, 0.8) .. (1, 0) .. 
    controls (0.2, -0.8) and (-0.2, -0.8) .. (-1, 0);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1.5, 0.75) -- (-0.75, 0) -- (-1.5, -0.75) ;
    \draw[thick] (0, 0) circle (0.75cm);
    \draw[thick, dashed] (0.53, 0.53) -- (1.25, 1.25);
    \draw[thick, dashed] (0.53, -0.53) -- (1.25, -1.25);
    \draw[thick] (0.53, -0.53) .. controls (0.2, -0.2) 
    and (0.2, 0.2) .. (0.53, 0.53);
\end{tikzpicture}
\ + \ 
\begin{tikzpicture}[scale=0.5, baseline={([yshift=-.5ex]current bounding box.center)}]
    \draw[thick, dashed] (-1.75, -0.75) -- (-1, 0) -- (-1.75, 0.75);
    \draw[thick, dashed] (1.75, -0.75) -- (1, 0) -- (1.75, 0.75);
    \draw[thick] (-1, 0) .. controls (-0.2, 0.8) and (0.2, 0.8) .. (1, 0) .. 
    controls (0.2, -0.8) and (-0.2, -0.8) .. (-1, 0);
\end{tikzpicture}
\end{split}
\end{displaymath}

\begin{equation}
\label{eq:f_functions}
f_1 = 
\end{equation}

\end{document}

@@

At the critical point it is convenient to introduce the dimensionless 
quantity
\begin{equation}
    \bar{\varphi} =
    \left(\frac{2\pi^2}{9} V g\right)^{\frac{1}{4}} \bar{\phi}\,,
\end{equation}
whose probability distribution is:
\begin{equation}
\label{eq:log_pdf_critical}
\begin{split}
    -\log p(\bar{\varphi}) = \mathrm{const.} +  
    &g^{\frac{1}{2}}\bar{\varphi}^2 
      \left(\frac{\pi}{\sqrt{2}}\left(f_1(0) - 1\right)  
      + O\left(g\right) \right) + \\
    &\bar{\varphi}^4\left(
      1 - \frac{g}{2}f_2\left(0\right)
      + O\left(g^2\right)\right) + \\
    &g^{\frac{3}{2}} \bar{\varphi}^6 \left(\frac{f_3(0)}{3\sqrt{2} \pi}
    + O\left(g\right)\right) + O\left(g^2\bar{\varphi}^8\right)\,.
\end{split}
\end{equation}

This expression has a residual dependence on system size via dimensional
transmutation of the renormalized coupling $g$, which at the critical point $r
= 0$ follows the simplified renormalization group equation
\begin{equation}
    \frac{1}{g(R_2)} - \frac{1}{g(R_1)} = \log\frac{R_2}{R_1}\,.
\end{equation}

Therefore, as system size grows at fixed bare couplings $r_0$ $u_0$, the
renormalized coupling $g$ goes to zero, and the all terms in
(\ref{eq:log_pdf_critical}) besides the quartic vanish. Thus we conclude
that, for sufficiently large system size: 
\begin{equation}
    -\log p(\phi) \sim \mathrm{const} + \frac{2\pi^2}{9} V g\bar{\phi}^4\,.
\end{equation}
